Uncovering invariant harmonic structures and schemata in music \\ \small A Deep Learning approach with contrastive learning and inductive bias

### 2023-03-04 introduction/Subsections/Introduction.tex
Updates from Overleaf

Music is a complex and fascinating art form studied by scholars and musicians for centuries. Music theory plays a crucial role in understanding the underlying principles of how music is organized and composed, and analyzing the structure of music is a challenging task.

Recently, Music Structure Analysis (MSA) has emerged as a subfield of Music Information Retrieval (MIR), aiming to develop computational methods to analyze the structure of music. This master thesis investigates the relationship between MSA, aural skills, and the underlying musical structure in musical composition.

In recent years, the utilization of neural networks has allowed for the creation of low-dimensional musical embeddings that capture important musical features and enable the computation of semantic similarity between musical passages. By learning an embedding function $f(x) \in \mathbb{R}^d$ that maps a given passage $x$ to a vector in the real space $\mathbb{R}^d$, researchers can quantify the semantic similarity between two musical passages using the Euclidean distance, which is represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. Adequate embedding space is essential in ensuring that the generated output aligns with human expectations, especially in interactive applications where a human performer influences the machine's response. The success of such systems depends on the embedding space's ability to interpret music in a way that aligns with human perception.

The thesis explores the musical elements that constitute music\footnote{The term "music" used in this context refers specifically to Western tonal music tradition. It is assumed that the musical structures and elements discussed are based on this tradition and may not apply to other musical styles or cultures.} structure analysis, including melody, harmony, rhythm, and form. Aural skills involve perceptual and cognitive abilities that enable individuals to comprehend musical works through listening. The underlying musical structure refers to the systematic arrangement of musical elements that results in a coherent and expressive musical work.

This thesis provides a comprehensive analysis of musical works from diverse styles and genres, utilizing computational techniques from MIR and incorporating a musicological perspective to contextualize the findings.

This research aims to offer fresh insights into these components of musical composition and advance our understanding of their relationships.

>>>>

The purpose of this work is to generalize the fundamental harmonic structure of music to facilitate its application in downstream tasks. Music, being a complex and fascinating art form, has been extensively studied by scholars and musicians for centuries. Understanding how music is organized and composed is critical, and music theory plays a vital role in this regard. Analyzing music structure poses a significant challenge as it comprises several elements, including melody, harmony, rhythm, and form.

In recent years, the utilization of neural networks has made it possible to create low-dimensional musical embeddings that capture crucial musical features and enable the computation of semantic similarity between musical passages. By acquiring an embedding function $f(x) \in \mathbb{R}^d$ that maps a given passage $x$ to a vector in the real space $\mathbb{R}^d$, researchers can measure the semantic similarity between two musical passages using the Euclidean distance, as represented by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The effectiveness of such systems depends on the embedding space's ability to interpret music in a way that aligns with human perception. Therefore, a suitable embedding space is crucial to ensure that the output aligns with human expectations.

The thesis under discussion explores the various musical elements that constitute music structure analysis, which includes melody, harmony, rhythm, and form. Aural skills are required to comprehend musical works through listening, as they involve perceptual and cognitive abilities. The underlying musical structure refers to the systematic arrangement of musical elements that results in a coherent and expressive musical work.

To provide a comprehensive analysis of musical works from diverse styles and genres, the thesis employs computational techniques from MIR and incorporates a musicological perspective to contextualize the findings. The objective is to offer fresh insights into these components of musical composition and advance our understanding of their relationships.

The study investigates the correlation between MSA, aural skills, and the underlying musical structure in musical composition. Through this research, the thesis aims to provide a better understanding of musical composition components, their relationships, and their practical applications.

### 2023-03-04 introduction/Subsections/Motivation.tex
Updates from Overleaf

As a music lover and connoisseur, I am interested in gaining a deeper understanding of the fundamental building of musical composition, the various techniques used to create musical works, and capturing underlying relationships among them. I am deeply interested in learning how to retrieve embedded information within music. The study of music structure analysis provides a means to systematically analyze and understand musical pieces. It has traditionally been approached through music theory, formal musical notation, musical analysis techniques, and machine learning, which I plan to blend.

With the advent of artificial intelligence and machine learning, Music Information Retrieval (MIR) has expanded to encompass musically motivated neural network architectures. These networks are designed to incorporate musical concepts and perceptual approaches and use this knowledge to analyze and understand the structure of musical pieces in a way that goes beyond traditional music analysis techniques.

As a professional musician and music transcriber, I am interested in understanding how deep neural networks (DNNs) filter musical features. In other words, I am eager to learn how DNNs listen~\cite{7500246} to music.

By studying music structure analysis using musically motivated neural networks, we will gain an in-depth understanding of the latest advancements in the field and how these techniques are being used to revolutionize our understanding of music. Additionally, we will be able to contribute to developing new and innovative methods for musical analysis~\cite{Huang2019MusicTG}

>>>>

As a music lover, professional musician, and connoisseur, I am interested in gaining a deeper understanding of the fundamental building of musical composition, the various techniques used to create musical works, and capturing underlying relationships among them. I am deeply interested in learning how to retrieve embedded information within music. The study of music structure analysis provides a means to systematically analyze and understand musical pieces. It has traditionally been approached through music theory, formal musical notation, musical analysis techniques, and machine learning, which I plan to blend.

With the advent of Artificial Intelligence (AI) and Machine Learning (ML), Music Information Retrieval (MIR) has expanded to encompass musically motivated neural network architectures. These networks are designed to incorporate musical concepts and perceptual approaches and use this knowledge to analyze and understand the structure of musical pieces in a way that goes beyond traditional music analysis techniques.

As a professional musician and music transcriber, I am interested in understanding how Deep Neural Networks (DNNs) filter musical features. In other words, I am eager to learn how DNNs listen to music.

By studying music structure analysis using musically motivated neural networks, we will gain an in-depth understanding of the latest advancements in the field and how these techniques are being used to revolutionize our understanding of music. Additionally, we will be able to contribute to developing new and innovative methods for musical analysis.

### 2023-03-04 music analysis/Equations/Novelty equation.tex
Updates from Overleaf

% N is the number of time frames, S_{ij} is the similarity between time frames i and j, S_{ii} is the similarity between time frame i and itself, and n_i is the novelty value for time frame i. The numerator of the expression represents the sum of similarities between time frame i and all other time frames, minus the average similarity across all time frames. The denominator of the expression is the standard deviation of the similarities, which is used to normalize the values to ensure that the novelty values are between -1 and 1.%

>>>>

Where:
\begin{itemize}
\item $N$ is the number of time frames.
\item $S_{ij}$ is the similarity between time frames $i$ and $j$.
\item $S_{ii}$ is the similarity between time frame $i$ and itself.
\item $n_i$ is the novelty value for time frame $i$.
\end{itemize}

The numerator of the expression represents the sum of similarities between time frame $i$ and all other time frames, minus the average similarity across all time frames. The denominator of the expression is the standard deviation of the similarities, which is used to normalize the values to ensure that the novelty values are between -1 and 1.

### 2023-03-04 music analysis/MSA.tex
Updates from Overleaf

In the following section, we will examine the state of MSA, including its challenges and potential for future advancements. Topics such as data representation, computational complexity, and evaluation metrics will be discussed to provide a nuanced summary of the field. Additionally, we will consider the impact of new trends and technological advancements on MSA and the directions in which the field is likely to evolve. Through this examination, we aim to provide an overview of MSA's technical and musical aspects and its challenges~\cite{Nieto2020}~\cite{Chaki2021}.

@@

As an art form, music is open to interpretation and can evoke different emotions and meanings in different listeners. This subjectivity makes it difficult to objectively analyze the structure of a piece of music, as different annotators may place boundaries between segments differently based on their perception and interpretation of the music. It can also have multiple structures, including hierarchical structures ~\cite{hiererchical}, that makes it difficult to untangle the underlying structure.

@@

@@

MSA has a long history, dating back to music theory and composition studies. In the 18th and 19th centuries, music theorists such as Rameau~\cite{christensen2004rameau} and Schenker~\cite{schenkerdocumentsonline} developed analytical frameworks for understanding the structure of music in terms of harmony, melody, and form.

@@

Heinrich Schenker~\cite{schenkerdocumentsonline} was a prominent music theorist who developed \textit{Schenkerian Analysis}, a method for analyzing tonal music. This method is based on the concept of "the Ursatz," a fundamental structure of three levels: foreground, middle-ground, and background. Schenker's approach emphasizes the horizontal dimension of music, specifically melody, and the relationships between individual notes and the piece's overall structure. His ideas have significantly impacted the field of music theory and analysis, and the method is still widely used today. While the approach is primarily theoretical and relies on human interpretation, it can also be approached through computer algorithms such as CNNs; the results are not the same as human interpretation.

@@

Schenkerian analysis uses specialized musical notation and is either presented in a generative manner, starting from the Ursatz and ending at the score or in a reductive manner, starting from the score and reducing it to its fundamental structure. The Ursatz is presented in an arrhythmic graph, and rhythmic signs display hierarchical relationships between events. \footnote{In Schenkerian analysis, structural levels refer to the representation of a piece of music at different levels of abstraction. These levels typically include the foreground, middle ground, and background. Schenker posits that musical form, as represented through these structural levels, can be understood as an energy transformation, specifically the transformation of forces that flow from the background to the foreground through the various levels.}

@@

Aural music skills are the capacity to comprehend and analyze musical compositions through listening. These skills include identifying musical elements such as rhythm, melody, harmony, timbre, and form, performing music from memory, and transcribing audio recordings into written musical scores.

In the context of music, having strong aural skills is crucial for achieving a complete understanding of the art form. It enables musicians to make knowledgeable decisions about their playing and interpretation, as well as to evaluate the playing of others critically. Additionally, aural skills are an integral part of music education, playing a role in deepening one's appreciation of music and promoting creativity and musical expression.

Aural skills can be honed through various activities, such as listening to music, singing, playing musical instruments, and analyzing music through written or visual means. Formal music education teaches aural skills through ear training, dictation, and sight-reading exercises.

Aural music skills are crucial not only for an individual's personal musical growth but also for gaining a deeper understanding of the underlying musical structure of a piece. Identifying and analyzing musical elements through listening allows one to comprehend their relationships and how they contribute to the overall musical structure. This ability to perceive and understand musical structure is fundamental to making informed decisions and interpretations.

In addition, solid aural skills enable musicians to communicate their musical ideas to others effectively, as well as to more fully comprehend the musical ideas of others. They provide a common ground for musical discussion and collaboration, allowing musicians to create and perform music at a higher level.

Moreover, aural skills also play an essential role in preserving and continuing musical traditions, as they allow musicians to accurately pass down and preserve the musical heritage of their culture. By recognizing and understanding the underlying musical structures of traditional music, musicians can maintain its authenticity and ensure its continuation for future generations.

In conclusion, aural music skills are crucial for personal musical growth and a deeper understanding of the musical structure, effective musical communication, collaboration, and preserving musical traditions.

>>>>

In the following section, we will examine the state of MSA, including its challenges and potential for future advancements. Topics such as data representation, computational complexity, and evaluation metrics will be discussed to provide a nuanced summary of the field. Additionally, we will consider the impact of new trends and technological advancements on MSA and the directions in which the field is likely to evolve. Through this examination, we aim to provide an overview of MSA's technical and musical aspects and its challenges.

@@

As an art form, music is open to interpretation and can evoke different emotions and meanings in different listeners. This subjectivity makes it difficult to objectively analyze the structure of a piece of music, as different annotators may place boundaries between segments differently based on their perception and interpretation of the music. It can also have multiple structures, including hierarchical structures, that make it difficult to untangle the underlying structure.

@@

The method of computing a novelty curve from a self-similarity matrix is widely used in Music Information Retrieval and audio analysis to assess the deviation from the average similarity of a given time frame in a music signal. It is primarily driven by Digital Signal Processing (DSP) and not necessarily motivated by musical considerations, yet it can be valuable in music segmentation, beat tracking, and structural analysis. However, the approach may be vulnerable to the choice of self-similarity measure used to calculate the matrix, and it may require further domain-specific knowledge and interpretation to extract practical information. Besides, the novelty curve could be limited in its ability to reflect musically meaningful events as it measures the deviation from the average similarity across time frames, rather than following musically motivated criteria. Therefore, the analysis and interpretation of the novelty curve may require additional caution to ensure its efficacy in capturing useful information about the music signal.

@@

MSA has a long history, dating back to music theory and composition studies. In the 18th and 19th centuries, music theorists such as Rameau and Schenker developed analytical frameworks for understanding the structure of music in terms of harmony, melody, and form.

@@

Heinrich Schenker was a prominent music theorist who developed \textit{Schenkerian Analysis}, a method for analyzing tonal music. This method is based on the concept of "the Ursatz," a fundamental structure of three levels: foreground, middle-ground, and background. Schenker's approach emphasizes the horizontal dimension of music, specifically melody, and the relationships between individual notes and the piece's overall structure. His ideas have significantly impacted the field of music theory and analysis, and the method is still widely used today. While the approach is primarily theoretical and relies on human interpretation, it can also be approached through computer algorithms such as CNNs; the results are not the same as human interpretation.

@@

Schenkerian analysis uses specialized musical notation and is either presented in a generative manner, starting from the Ursatz and ending at the score, or in a reductive manner, starting from the score and reducing it to its fundamental structure. The Ursatz is presented in an arrhythmic graph, and rhythmic signs display hierarchical relationships between events. \footnote{In Schenkerian analysis, structural levels refer to the representation of a piece of music at different levels of abstraction. These levels typically include the foreground, middle ground, and background. Schenker posits that musical form, as represented through these structural levels, can be understood as an energy transformation, specifically the transformation of forces that flow from the background to the foreground through the various levels.}

@@

The capacity to comprehend and analyze musical compositions through listening requires a range of abilities. These include recognizing musical elements such as rhythm, melody, harmony, timbre, and form, performing music from memory, and transcribing audio recordings into written musical scores.

Having a keen sense of hearing is crucial for a complete understanding of music. It enables musicians to make informed decisions about their playing and interpretation, as well as critically evaluate the playing of others. Aural abilities are essential in music education, deepening appreciation of music and promoting creativity and expression.

Developing these skills can be achieved through various activities, such as attentive listening, singing, playing instruments, and analyzing music through written or visual means. Formal music education includes ear training, dictation, and sight-reading exercises.

Strong aural abilities facilitate personal musical growth and understanding of the underlying musical structure of a piece. Identifying and analyzing musical elements through listening allows comprehension of their relationships and contribution to the overall musical structure. This is fundamental to making informed decisions and interpretations.

Solid listening skills enable effective communication and collaboration among musicians. They provide a common ground for musical discussion, creating and performing music at a higher level.

The ability to perceive and understand musical structure is also essential for preserving musical traditions by accurately passing down the musical heritage of a culture. Recognizing and understanding the underlying musical structures of traditional music maintains its authenticity and ensures its continuation for future generations.

In conclusion, the ability to comprehend and analyze musical compositions through listening is vital for personal musical growth, understanding the musical structure, effective communication, collaboration, and preserving musical traditions.
